---
title: hierr:\ An R Package for Hierarchical Regularized Regression
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "readme_files/readme"
)
```

[![Build Status](https://travis-ci.org/USCbiostats/hierr.svg?branch=master)](https://travis-ci.org/USCbiostats/hierr)
[![codecov](https://codecov.io/gh/USCbiostats/hierr/branch/master/graph/badge.svg)](https://codecov.io/gh/USCbiostats/hierr)

# Setup

1. For Windows users, install [RTools](https://cran.r-project.org/bin/windows/Rtools/) (not an R package)
2. Install the R package [devtools](https://github.com/hadley/devtools)
2. Install hierr package with the install_github function (optionally you can install the most recent / potentially unstable development branch)
3. Load the package 

```{r, eval = FALSE}
library(devtools)

# Master branch
install_github("gmweaver/hierr")

# Or the development branch
install_github("gmweaver/hierr", ref = "development")
```
```{r load package}
library(hierr)
```

# A First Example

As an example of how you might use the hierr package, we have provided a small set of simulated external data variables (ext), predictors (x), and a continuous outcome variable (y). First, load the example data:

```{r load data}
x <- hierr::x
y <- hierr::y
ext <- hierr::ext
```

#### Fitting a Model

To fit a linear hierarchical regularized regression model, you can use the main `hierr` function. At a minimum, you must specify the predictor matrix (x), outcome variable (y), external data matrix (ext), and outcome distribution. By default, a ridge penalty is applied to the predictors and a lasso penalty is applied to the external data.

```{r fit simple model}
hierr_model <- hierr(x = x, y = y, external = ext, family = "gaussian")
```

#### Modifying Regularization Terms

To modify the regularization terms and penalty path associated with the predictors or external data, you can use the `definePenalty` function. This function allows you to configure the following regularization attributes:

* Regularization type 
    - Ridge = 0
    - Elastic Net = (0, 1)
    - Lasso / Quantile = 1 (quantile only on development branch currently) 
* Penalty path
    - Number of penalty values in the full penalty path (default = 20)
    - Ratio of min(penalty) / max(penalty) 
* User-defined set of penalties

As an example, we may want to apply a ridge penalty to both the x variables and external data variables. In addition, we may want to have 30 penalty values computed for the regularization path associated with both x and external. We modify our model fitting as follows.

```{r fit simple model with other penalty}
myPenalty <- definePenalty(penalty_type = 0, penalty_type_ext = 0, num_penalty = 30, num_penalty_ext = 30)
hierr_model <- hierr(x = x, y = y, external = ext, family = "gaussian", penalty = myPenalty)
```

#### Tuning Penalty Parameters by Cross-Validation

In general, we need a method to determine the penalty values that produce the optimal out-of-sample prediction. We provide a simple two-dimensional grid search that uses k-fold cross-validation to determine the optimal values for the penalties. The cross-validation function `cvhierr` is used as follows.

```{r, include = FALSE}
set.seed(123)
```

```{r fit k-fold cv}
cv_hierr <- cvhierr(x = x, y = y, external = ext, family = "gaussian")
```
To visualize the results of the cross-validation we provide a contour plot of the mean cross-validation error across the grid of penalties with the `plot` function.

```{r cv_results}
plot(cv_hierr)
```
